---
title: "Analysis on Factors Impacting Internet Usage"
output: html_document
author: Krupa Patel
---

Content List:
-> Title
-> Importing the Data
-> Merging the Data
-> Cleaning the Data
-> Mapping the Data
-> Problem Introduction
-> Problem #1: Does internet access decrease spending on education due to an increase in productivity?
-> Problem #2: Does internet access decrease spending on health due to an increase in productivity?
-> Problem #3: How does the democracy score compare with internet usage? 
-> Conclusions
-> Recommendation
-> Appendix A (Roles Each Team Member Played in the Project) 


---------------------------------------------------------------------------------------------------------------------          
Importing the Data:

manually imported the data using the read_csv()

```{r}
library(readr)
life_expect <- read_csv("lifeexpect_cia2017.csv")
gdpppp <- read_csv("gdpppp_cia2017.csv")
health_expend <- read_csv("healthexpend_cia2017.csv")
democracy_index <- read_csv("democracyindex2017.csv")
education <- read_csv("education_cia2017.csv")
internet_users <- read_csv("internetusers_cia2017.csv")
population <- read_csv("population_cia2017.csv")
world_regions <- read_csv("world_regions.csv")
telephonelines <- read_csv("telephonelines_cia2017.csv")
```
---------------------------------------------------------------------------------------------------------------------
Merging The Data:

By using the function full_join(), combined each of the datasets into one data set by using country as a key.

```{r}
library(tidyverse)
dirty_full_data <- full_join(internet_users, population, by="Country")
dirty_fuller_data <- full_join(dirty_full_data, telephonelines, by="Country")
dirty_fullest_data <- full_join(dirty_fuller_data, world_regions, by="Country")

d1 <- full_join(dirty_fullest_data,gdpppp,by="Country")
d2<- full_join(d1,health_expend,by="Country")
names(d2)[17] <- "GDPHealth"      # changed the name "% of Health to "GDPHealth
d3<- full_join(d2,education,by="Country")
d4<-full_join(d3,democracy_index,by="Country")
d5<-full_join(d4,life_expect,by="Country")
names(d5)[20] <- "GDPEducation"  # changed the name "% of Education to "GDPEducation
d6 <- select(d5,-"Date of Information.x",-"Rank.y",-"Date of Information.y",-"Rank.x.x",-"Date of Information.x.x",-"Rank.y.y",-"Rank.x.x.x",-"Date of Information.y.y",-"Date of Information.x.x.x",-"Rank.y.y.y",-"Date of Information.y.y.y",-"Rank.x.x.x.x",-"Rank.y.y.y.y",-"Date of Information",-"Rank.x")

names(d6)[2] <- "Internet_users"                   #"INTERNET USER"to "Internet_user"
names(d6)[3] <- "Population"                       #"POPULATION" to "Population"
names(d6)[4] <- "Telephone_MainLines"              #"TELEPHONES - MAIN LINES IN USE" to "Telephone_MainLines"
names(d6)[7] <- "GDP_Per_Capita"                   #"GDP - PER CAPITA (PPP)" to "GDP_Per_Capita"
names(d6)[10] <- "Democracy_Score"                 #"Score" to "Democracy_Score"
names(d6)[11] <- "Electoral_Process_Pluralism"     #"Electoral processand pluralism" to "Electoral_Process_Pluralism"
names(d6)[17] <- "Life_Expectancy"                 #"Years" to "Life_Experctancy"
```
---------------------------------------------------------------------------------------------------------------------
Cleaning the Data:
 changed the name from Asic & Pacific to Asia & Pacific

```{r}
d6$Region[d6$Region == "Asic & Pacific"] <- "Asia & Pacific"
```

defining internet usage: internet usage is the proportion of population to the internet users in that country.  

```{r}
d7<- mutate(d6, Prop_usage = ((Internet_users)/ Population))
```

removed the dollar sign from data set to ensure that the GDP per Capita values are integer values rather that character values.

```{r}
d7$GDP_Per_Capita_2 = as.numeric(gsub("[\\$,]","",d7$GDP_Per_Capita))
```
---------------------------------------------------------------------------------------------------------------------
Mapping the Data:

```{r}

d7[7,5] = "Europe"
d7[4,1] = "USA"
d7[11,1] = "UK"
```

```{r}
#install.packages("maps")
library(maps)
gg2 <- ggplot()
gg2 <- gg2 + geom_map(
  data = world,
  map = world,
  aes(x = long, y = lat, map_id = region),
  fill = "#ffffff",
  color = "#ffffff",
  size = 0.20
)
  gg2 <- gg2 + scale_fill_continuous(low = 'thistle2', high = 'darkblue',
  guide = 'colorbar')
gg2 <- gg2 + geom_map(
  data = d7_map,
  map = world,
  aes(fill = d7_map$GDP_Per_Capita_2, map_id = region),
  color = "#ffffff",
  size = 0.15
  )
  
  gg2 <- gg2 + scale_fill_continuous(low = 'thistle2', high = 'darkblue',
  guide = 'colorbar')
  
gg2
```

```{r}
world <- map_data("world") 
d7_map <- d7 %>% rename(region = Country) 

d7_map <- semi_join(d7_map, world, by = "region")
d7_map
gg <- ggplot()
gg <- gg + geom_map(
  data = world,
  map = world,
  aes(x = long, y = lat, map_id = region),
  fill = "#ffffff",
  color = "#ffffff",
  size = 0.20
)

gg <- gg + geom_map(
  data = d7_map,
  map = world,
  aes(fill = d7_map$Prop_usage, map_id = region),
  color = "#ffffff",
  size = 0.15
  )
  
  gg <- gg + scale_fill_continuous(low = 'thistle2', high = 'darkblue',
  guide = 'colorbar')
gg
```


```{r}   
d7 <- filter(d7, Region != "NA")            #removes NA values in region
d7 <- filter(d7, Region != "Unknown")       #removes unknown values in Region
ggplot(data=d7)+geom_boxplot(mapping=aes(x=Region,y=Prop_usage))
```
---------------------------------------------------------------------------------------------------------------------
Problem Introduction:

Problem #1: Does internet access decrease spending on education due to an increase in productivity?
Problem #2: Does internet access decrease spending on health due to an increase in productivity?
Problem #3: How does the democracy score compare with internet usage?

---------------------------------------------------------------------------------------------------------------------
Problem #1: Does internet access decrease spending on education due to an increase in productivity?


```{r}
#created a new data set Edu_Internet made up of Country,GDPEducation,Prop_Usage,Region and "Global South" as the variables
Edu_Internet <- select(d7,Country,GDPEducation,Prop_usage,Region,"Global South")

#Removes the NA values in the GDPEducation Variables 
Edu_Internet <-filter(Edu_Internet,GDPEducation != "NA")

#Removes the NA values in the Prop_Usage Variables 
Edu_Internet <-filter(Edu_Internet,Prop_usage != "NA")

#Removes the NA values in the Region Variables
Edu_Internet <-filter(Edu_Internet,Region != "NA")

#Removes the NA values in the "Global South" Variables
Edu_Internet <-filter(Edu_Internet,"Global South" != "NA")

#Changes the the Global South value at the Edu_Internet[5,5] from "NA" to Global North for the country of Russia 
Edu_Internet[5,5] = "Global North"

#checks that ithe number of NA values sums up to 0 since all NA values have been remoeved
sum(is.na(Edu_Internet))
```


```{r}
library(modelr)

#Creates a scatter plot of the Prop_Usage(Internet Usage/Population) against the GDPeducation and the n facet-wrapped them according to the regions
ggplot(data=Edu_Internet)+geom_point(mapping=aes(x=Prop_usage,y=GDPEducation))+facet_wrap(~ Region, nrow = 3)
```

```{r}
# Creates a scatter plot of the Prop_usage and the GDPEducation and color code them based on the region.
ggplot(data=Edu_Internet)+geom_point(mapping=aes(x=Prop_usage,y=GDPEducation,colour=Edu_Internet$"Global South"))
```


```{r}
#function for the linear model
mod_Edu_Internet <- lm(GDPEducation ~ Prop_usage, data = Edu_Internet)

#The code below was used to generate a prediction model
grid <- Edu_Internet %>%
  data_grid(Prop_usage, .model = mod_Edu_Internet) %>%
  add_predictions(mod_Edu_Internet)
grid
ggplot(grid, aes(Prop_usage, pred)) + geom_point()
```

--------------------------------------------------------------------------------------------------------------------- 
Problem #2: Does internet access decrease spending on health due to an increase in productivity?


```{r}
Health_Internet <- select(d7,Country,GDPHealth,Prop_usage,Region,"Global South",Life_Expectancy)
na_remove<-which(!complete.cases(Health_Internet))
Health_Internet <- Health_Internet[-na_remove,]
sum(is.na(Health_Internet))
```

```{r}
library(modelr)

ggplot(Health_Internet, mapping = aes(x=Prop_usage,y=Life_Expectancy))+geom_point() + geom_smooth(se = FALSE) 
mod_life_expectancy<-lm(Life_Expectancy~Prop_usage,data=Health_Internet)
grid<- Health_Internet %>% 
  data_grid(Prop_usage = seq_range(Prop_usage,40)) %>%
  add_predictions(mod_life_expectancy,"Life_Expectancy")
```

```{r}
ggplot(Health_Internet,aes(Prop_usage,Life_Expectancy))+
  geom_hex(bins = 50)+
  geom_line(data=grid,color = "red",size = 1)+
  geom_point()
  
```

```{r}
Health_Internet <- Health_Internet %>%
  add_residuals(mod_life_expectancy)
ggplot(Health_Internet, aes(Prop_usage, resid)) +
  geom_ref_line(h=0) +
  geom_line()
lm(Health_Internet$Life_Expectancy ~ Health_Internet$Prop_usage)$coeff[2]
```

```{r}
mod_life_expectancy_GDP<-lm(Life_Expectancy~GDP_Per_Capita_2,data=d7)
grid<- d7 %>% 
  data_grid(GDP_Per_Capita_2 = seq_range(GDP_Per_Capita_2,40)) %>%
  add_predictions(mod_life_expectancy_GDP,"Life_Expectancy")
ggplot(d7,aes(x=log(GDP_Per_Capita_2), y=Life_Expectancy)) + geom_point() +
  geom_hex(bins = 50)+
  geom_line(data=grid,color = "red",size = 1)
```

---------------------------------------------------------------------------------------------------------------------          
Problem #3: How does the democracy score compare with internet usage? 


```{r}

Democracy_Internet <- select(d7,Country,Democracy_Score,Prop_usage,Region,"Global South", Category, GDP_Per_Capita_2)
na_remove<-which(!complete.cases(Democracy_Internet))
Democracy_Internet <- Democracy_Internet[-na_remove,]
#View(Democracy_Internet)

```

```{r}
gg2 <- ggplot(data = Democracy_Internet, mapping = aes(x = reorder(Category,Prop_usage,FUN = median ), y = Prop_usage))
gg2 + geom_boxplot()

ggplot(data = Democracy_Internet, mapping = aes(x = reorder(Category,log(GDP_Per_Capita_2),FUN = median), y = log(GDP_Per_Capita_2))) + geom_boxplot()

```

---------------------------------------------------------------------------------------------------------------------          
Conclusions:

The proportion of internet users in a country is a strong indicator of how developepd a country is. Whether you are looking at it through a economical standpoint, health standpoint, or political standpoint, we see that the better the conditions are in the country, the more likely the country has a higher proportion of internet users. Although, it is clear that proportion of internet users does not have a clear relationship on how much a country spends on GDP, other than a slight relationship European countries' internet usage proportion and their education spending, how much a country spends on GDP is nevertheless an interesting case.
---------------------------------------------------------------------------------------------------------------------          



        


